### 2015-11-16 docs/NEMO/psyclone_in_nemo.tex
Add draft report/work plan from meeting at Met Office on 12-13/11/2015

subroutine ldf_lsp_triad(...)

  call invoke( horiz_density_grad(expan_coeff, temp, salin, density_grad) )
  if (partial_steps) then
    call invoke( bottom_grad_corr(density_grad, expan_coeff) )
  end if
  call invoke( vertical_density_grad(temp, salin, ...),       &
               recip_depth_w_mix(...),                        &
               set(..., 0),                                   &
               triad_slope_at_mixed_layer(density_grad, ...), &
               calc_triad(...) )

end subroutine ldf_lsp_triad