### 2016-01-22 docs/papers/roofline_model/gocean_roofline.tex
roofline: fleshing out text now we have results

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage{url}

\newcommand{\psykal}{{PS}y{KA}l}

\begin{document}

\title{Analysing the Performance of Shallow-Water Models with the
 Roofline Model}

\author{A.~.R.~Porter and R.~W.~Ford}

\maketitle

\section{Introduction}

We want to use the Roofline Model to establish the efficiency (or
otherwise) of two shallow-water benchmark codes that we have used in
the GOcean project. The first of these codes, Shallow, was originally
written by Paul Swarztrauber over 30 years ago and has been optimised
by various people over that time. In contrast, nemolite2d is a new,
un-optimised code that has been extracted from NEMO.

Although we have investigated how the performance of the \psykal
version of nemolite2d compares with that of the original, we have not
addressed how efficient the original actually is. In order to do so we
use the Roofline Model~\cite{roofline} which provides a relatively
simple way of characterising the performance of a code in terms of
whether it is memory-bandwidth bound or compute bound.

In this report we investigate the application of the Roofline Model to
these two, serial, shallow-water codes.

\section{Hardware Details}

We have performed this investigation on two different Intel CPUs; an
E5-1620 v.2 and an E5-2697 v.2. Both of these use the Intel Ivy Bridge
microarchitecture and have Ivy Bridge-EP cores. More details are given
in Table~\ref{TAB_stream_linpack}.

\begin{table}
\begin{tabular}{|l|c|c|c|c|}
\hline
  \multicolumn{5}{|c|}{E5-1620 v.~2, 3.7GHz} \\
  \hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{28.32} \\
\hline
 &  L1 Cache & L2 Cache & L3 Cache & Main memory \\
 &   32 KB   & 256 KB   & 10 MB    &  --         \\
\hline
STREAM2 (GB/s)   &  117  &  65.8  & 46.7 &  20.5 \\
  \hline
\multicolumn{5}{|c|}{E5-2697 v.~2, 2.7GHz}  \\
\hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{25.41} \\
\hline
 & L1 Cache & L2 Cache & L3 Cache & Main memory \\
 &   32 KB  & 256 KB   & 30 MB    &  --         \\
\hline
STREAM2 (GB/s) & 108 & 54.0 & 37.7  &  17.6 \\
\hline
\end{tabular}
\caption{The STREAM2 (DAXPY) and LINPACK benchmark results for a
  single thread running on each of the two CPUs considered here.}
\label{TAB_stream_linpack}
\end{table}

In applying the roofline model we follow the approach suggested
in~\cite{para_pearls} and use the well-known STREAM~\cite{stream} and
LINPACK~\cite{linpack} benchmarks in order to get the upper bounds on
the memory bandwidth and FLOPS for the test hardware. Since we are
using Intel CPUs we used the Intel Math Kernel Library implementation
of LINPACK.

Interpreting the results of the STREAM benchmark in order to get
bandwidth figures for all levels of the cache hierarchy is not
straightforward. We therefore used the STREAM2 benchmark which is
intended for this purpose. In Figure~\ref{FIG_stream2} we plot the
results obtained from the DAXPY component of the STREAM2 benchmark on
both of the CPUs examined here. The plateaux in bandwidth for L2, L3
and main memory are clear but that for L1 is considerably noisier. The
final values we have used in the roofline plots are given in
Table~\ref{TAB_stream_linpack}.

\begin{figure}
\includegraphics{stream2_desktop}
\includegraphics{stream2_archer}
\caption{Results from running the STREAM2 benchmark on the E5-1620
  (top) and E5-2697 (bottom) CPUs.}
\label{FIG_stream2}
\end{figure}

\section{Hardware Performance Counters}

Initially, we hoped to use Hardware Performance Counters (HWPCs) in
order to get an accurate measure of the FLOPS and memory traffic
associated with specific kernels in each of the models.

The likwid tool (\url{https://github.com/RRZE-HPC/likwid}) gives
access to HWPCs on Intel and AMD CPUs.
However, there are issues with the accuracy of these counters,
especially for FLOP-related measurements on Ivy Bridge architectures. A
tool is included in the likwid distribution that tests the accuracy of
HWPC-derived data by running micro-benchmarks for which the FLOP count
and memory traffic are known:
\url{https://code.google.com/p/likwid/wiki/IvyBridgeVerification}.
The results of doing this on an Intel Xeon E5-1620 v2 CPU are shown in
Figure~\ref{FIG_flops_dp_test}. It appears that once the memory
footprint spills out of L1 cache (which all of our nemolite2d
benchmark configurations do), the performance counter results
significantly overestimate the FLOPs performed.

\begin{figure}
\centering
\includegraphics[width=120mm]{FLOPS_DP_stream_ivybridgeEP.pdf}
\includegraphics[width=120mm]{FLOPS_DP_triad_ivybridgeEP.pdf}
\caption{Results of the likwid accuracy test for the FLOPS\_DP group
  and the stream (top) and triad (bottom) benchmark on an Intel Xeon
  E5-1620 v2.}
\label{FIG_flops_dp_test}
\end{figure}

We therefore need some other way of measuring FLOPS. We could use
likwid on an older Intel CPU (to which we have root access) or try
another tool on the IvybridgeEP system. Other possible tools include
Intel's VTune or OProfile.

Despite the unreliability of FLOP counts, we hope that the trends in
HWPC data as the problem size is varied will be useful. This raw data
is given in Table~\ref{TAB_hwpc_mom_u}.

\begin{table}
\begin{tabular}{l|l|l|l|l}
\hline
\hline
               & \multicolumn{4}{c}{Problem size} \\
               & 32     &    64        &    128       &   256 \\
\hline
Footprint (MB) & 0.1 & 0.5          & 2.3          & 9.2 \\
\hline
Runtime (RDTSC) [s]   & 0.144868 &  0.4485108   &   1.697142   &   6.446171  \\
Runtime unhalted [s]  & 1.199456e-01 & 4.475135e-01 & 1.750448e+00 &   6.734944e+00 \\
CPI                   & 6.431867e-01 & 6.423220e-01 & 6.503796e-01 &   6.348858e-01 \\
\hline
MFLOPS           & 3.112780e+03 & 3.877720e+03  & 4.163209e+03 & 4.304728e+03 \\
AVX MFLOPS       & 2.840625e+03 & 3.660086e+03  & 3.993397e+03 & 4.164647e+03 \\
Packed MUOPS/s   & 7.423377e+02 & 9.584935e+02  & 1.046781e+03 & 1.092280e+03 \\
Scalar MUOPS/s   & 2.077927e+02 & 1.306897e+02  & 7.294811e+01 & 3.784506e+01 \\
\hline
L1 DTLB load misses          &      2007     &      17015   &    1317912   &    6748712   \\
L1 DTLB load miss rate       &  2.949686e-06 &  6.607264e-06& 1.326630e-04 & 1.723565e-04 \\
L1 DTLB load miss durn [Cyc] &  3.970154e+01 &  3.080153e+01& 2.101039e+01 & 2.013584e+01 \\
L1 DTLB store misses         &       77      &       16     &     87257    &    401036    \\
L1 DTLB store miss rate      &  1.131668e-07 &  6.213119e-09& 8.783423e-06 & 1.024213e-05 \\
L1 DTLB store miss durn [Cyc]&  5.633766e+01 &       29     & 2.151055e+01 & 2.669793e+01 \\
\hline
L2D load bandwidth [MB/s]  &  3.811099e+03 &  5.030691e+03 & 1.080936e+04 & 1.446214e+04 \\
L2D load data volume [GB]  &   0.541220288 &   2.34655488  &  18.1215488  &   92.177536  \\
L2D evict bandwidth [MB/s] &  3.131102e+02 &  3.094459e+02 & 4.501487e+02 & 5.065637e+02 \\
L2D evict data volume [GB] &   0.04446528  &   0.144340352 &  0.75465984  &  3.22869248  \\
L2 bandwidth [MB/s]        &  4.600834e+03 &  5.493404e+03 & 1.130282e+04 & 1.498363e+04 \\
L2 data volume [GB]        &   0.653371776 &   2.562386304 & 18.948807808 & 95.501414592 \\
\hline
L3 load bandwidth [MB/s]  &  2.434255e+03 & 4.626177e+03 & 4.511686e+03 & 4.607617e+03 \\
L3 load data volume [GB]  &   0.371968512 &  2.07489024  &   7.6168768  &   29.692832  \\
L3 evict bandwidth [MB/s] &  5.652167e+02 & 4.828169e+02 & 2.771148e+02 & 2.722823e+02 \\
L3 evict data volume [GB] &   0.086368448 &  0.216548608 &  0.467840512 &  1.75466688  \\
L3 bandwidth [MB/s]       &  2.999472e+03 & 5.108994e+03 & 4.788801e+03 & 4.879899e+03 \\
L3 data volume [GB]       &   0.45833696  &  2.291438848 &  8.084717312 &  31.44749888 \\
\hline
\hline
\end{tabular}
\caption{Hardware performance counter data obtained using the likwid
  tool for the u-Momentum kernel on the Intel E5-1620 v2 processor.}
\label{TAB_hwpc_mom_u}
\end{table}

\section{Performance Analysis with the Roofline Model}

\subsection{Calculating Kernel Characteristics}

Although the Continuity kernel is very simple, the u-Momentum kernel
is not and in particular, includes a call to the {\tt sin()} function.
In order to estimate the cost of this function in FLOPs, the following
code was compiled (with the Intel compiler and -O1) and run with
likwid-perfctr:
\begin{verbatim}
  my_sum = 0.0d0
  do i = 1, 1000
    my_sum = my_sum + sin(real(i))
  end do
  write (*,*) 'My sum = ', my_sum
\end{verbatim}
This measurement revealed that the {\tt sin()} function requires
approximately 13 FLOPs.

A key component of the Roofline model is the Arithmetic Intensity of
the code being executed:
\begin{equation}
AI = \frac{\textrm{No. of floating-point operations}}{\textrm{Bytes fetched from memory}}
\end{equation}
Since HWPC data proved to be unreliable, we calculated this quantity
manually by examining the source code and counting the number of
memory references and arithmetic operations that it contained. In
doing this counting we assume that any references to adjacent array
elements are fetched in a single cache-line and thus only count once.

To illustrate this, we consider the first loop-nest in the
time-stepping section of the Shallow code. This is a doubly-nested
loop over {\tt i} and {\tt j} with {\tt i} innermost. The body of the
loop is:
\begin{verbatim}
CU(I+1,J) = .5d0*(P(I+1,J)+P(I,J))*U(I+1,J)
CV(I,J+1) = .5d0*(P(I,J+1)+P(I,J))*V(I,J+1)
Z(I+1,J+1) =(FSDX*(V(I+1,J+1)-V(I,J+1))- &
             FSDY*(U(I+1,J+1)-U(I+1,J)))/&
            (P(I,J)+P(I+1,J)+            &
             P(I+1,J+1)+P(I,J+1))
H(I,J) = P(I,J)+0.25d0*                   &
         (U(I+1,J)*U(I+1,J)+U(I,J)*U(I,J) & 
         +V(I,J+1)*V(I,J+1)+V(I,J)*V(I,J))
\end{verbatim}
This code writes to four memory locations ({\tt CU}, {\tt CV}, {\tt Z}
and {\tt H}) and reads from ten distinct locations.  However, if we
assume caching of reads from locations adjacent in memory ({\it e.g.} {\tt
  P(i,j)} and {\tt P(i+1,j)}) then we have only six distinct reads.
If we assume that each of these ten memory accesses will touch a
separate cache line then this code fragment will result in the moving
of ten cache lines, each of which is 64 bytes (eight double-precision
words). However, as noted above, this code fragment is the body of a
loop nest. Therefore, upon the next loop trip {\tt i} will have been
incremented by one and the majority of the accessed values will still
be in cache. In fact, a new cache line will only have to be fetched
every seven trips of the (innermost) loop and thus the mean memory
traffic per loop trip is approximately $640/7$ bytes.

We can also count the number of floating-point arithmetic operations
(FLOPs) in this code fragment which gives us 12 multiplications and 12
additions. The $AI$ for this code fragment is then $24*7/640 = 0.26$
FLOPs/byte.  If we do the same analysis for the (longer) u-Momentum
kernel of nemolite2d then we find that there are 51 addition
operations, 56 multiplication/division operations and one call to {\tt
  sin()}. There are 31 distinct memory references. This gives an $AI$
of approximately $(107+13)*7/(31*64)$ or 0.42 FLOPs/byte (where we
assume the call to {\tt sin()} costs 13 FLOPs). The results of this
analysis are summarised in Table~\ref{TAB_kernel_details}.

\begin{table}
\begin{tabular}{|l|c|c|c|}
\hline
Code                       & Shallow & \multicolumn{2}{c|}{nemolite2d} \\
\hline
Kernel                     & 1st loop nest & Continuity & u-Momentum \\
\hline                                
FLOPs                      & 24      &    14      &   107+13   \\
Distinct memory refs       & 10      &    10      &   31       \\
$AI$ (FLOPs/byte)          & 0.26    &  0.153     &   0.42     \\
Working-set size (bytes)   & $56n^2$ &  $64n^2$   &   $136n^2$ \\
\hline
\end{tabular}
\caption{Details of the kernels investigated with the roofline
  model. $n$ is the problem size -- the length of one side of the
  square domain.}
\label{TAB_kernel_details}
\end{table}

We measure the mean time to execute the u-Momentum kernel over the
whole domain as 0.0020 seconds (Intel v.16 on the E5-1620) for the
$256^2$ problem size. This gives us a performance of
$\frac{108*256^2}{0.002} = 3.5$ GFLOPS. In contrast, instrumenting
this section of code with the LIKWID marker API and using
likwid-perfctr we measure 4.7 GFLOPS.

For the $128^2$ domain, the measured time is 0.570033E-03 s. This
gives a performance of $108*128^2/0.570e-3 = 3.1$ GFLOPS which is 25\%
less than we measure directly with likwid.

Using these figures we can finally construct the Roofline model for
the two CPUs and the various kernels. In
Figure~\ref{FIG_roofline_archer} we plot the model for the E5-2697 CPU
and show the performance of the 1st loop-nest of Shallow and the
u-Momentum kernel of nemolite2d.

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_archer}
\caption{Comparison of the performance achieved by nemolite2d and
  Shallow on the E5-2697}
\label{FIG_roofline_archer}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_haswell}
\caption{Comparison of the performance achieved by different nemolite2d and
kernels on the E5-1620.}
\label{FIG_roofline_haswell}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=120mm]{nemolite2d_kernel_perf}
  \caption{Performance of the Continuity and u-Momentum kernels as a
    function of SIMD vectorisation and problem size. Results are for
    v.16 of the Intel compiler on a Xeon E5-1620 v.2 CPU.}
  \label{FIG_kernel_perf}
\end{figure}

\section{Tools}

This work has deliberately focused on the application of a relatively
simple performance model to two very simple, stencil based simulation
codes.  Despite this, getting the data necessary for the construction
of the roofline model has not been straightforward. Once the roofline
model had been constructed, it was also difficult to establish why the
performance of a section of code was apparently poor.

The reasons for this include:
\begin{itemize}
\item HWPC data on its own doesn't give much insight into performance issues;
\item FLOPS data produced by likwid is unreliable on Ivy Bridge processors;
\item Persuading any tool to give HWPC data can be difficult. likwid
  is relatively straightforward but root access was required on a
  Redhat-based linux distribution. Thus far we haven't succeeded in
  getting this data from Intel's VTune.
\item The Intel compiler tends to report ``Loop was vectorised'' but
  it is hard to judge how efficient this vectorisation is. For
  instance we have found occasions where backing-off from AVX to SSE
  instructions actually gave better performing code. Ways of
  establishing the quality of the code produced by the compiler are
  missing;
\end{itemize}

It feels as though there is a need for a tool that addresses some of
these issues. It could be that Intel's Vectorization Advisor will be
useful and this requires investigation.

\section{Conclusions}

\bibliographystyle{unsrt}
\bibliography{gocean_roofline}

\end{document}