### 2016-03-07 docs/papers/roofline_model/gocean_roofline.tex
roofline: add more discussion of Continuity SIMD results.

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage{url}

\newcommand{\psykal}{{PS}y{KA}l}

\begin{document}

\title{Analysing the Performance of Shallow-Water Models with the
 Roofline Model}

\author{A.~.R.~Porter and R.~W.~Ford}

\maketitle

\section{Introduction}

We want to use the Roofline Model to establish the efficiency (or
otherwise) of two shallow-water benchmark codes that we have used in
the GOcean project. The first of these codes, Shallow, was originally
written by Paul Swarztrauber over 30 years ago and has been optimised
by various people over that time. In contrast, nemolite2d is a new,
un-optimised code that has been extracted from NEMO.

Although we have investigated how the performance of the \psykal
version of nemolite2d compares with that of the original, we have not
addressed how efficient the original actually is. In order to do so we
use the Roofline Model~\cite{roofline} which provides a relatively
simple way of characterising the performance of a code in terms of
whether it is memory-bandwidth bound or compute bound.

In this report we investigate the application of the Roofline Model to
these two, serial, shallow-water codes.

\section{Hardware Details}

We have performed this investigation on two different Intel CPUs; an
E5-1620 v.2 and an E5-2697 v.2. Both of these use the Intel Ivy Bridge
microarchitecture and have Ivy Bridge-EP cores. More details are given
in Table~\ref{TAB_stream_linpack}.

\begin{table}
\begin{tabular}{|l|c|c|c|c|}
\hline
  \multicolumn{5}{|c|}{E5-1620 v.~2, 3.7GHz} \\
  \hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{28.32} \\
\hline
&  L1 Cache & L2 Cache & L3 Cache & Main memory \\
$n$-way set associative & 8    & 8   & 10 & -- \\
Size                  &   32 KB/core   & 256 KB/core   & 10 MB    &  -- \\
\hline
STREAM2 (GB/s)   &  160  &  61.0  & 46.0 &  20.5 \\
  \hline
\multicolumn{5}{|c|}{E5-2697 v.~2, 2.7GHz}  \\
\hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{25.41} \\
\hline
               & L1 Cache & L2 Cache & L3 Cache & Main memory \\
$n$-way set associative & 8    & 8   & 20 & -- \\
Size           &   32 KB/core  & 256 KB/core   & 30 MB    &  -- \\
\hline
STREAM2 (GB/s) & 132      & 59.0     & 40.0     &  20.2 \\
\hline
\end{tabular}
\caption{The STREAM2 (DAXPY) and LINPACK benchmark results for a
  single thread running on each of the two CPUs considered here.}
\label{TAB_stream_linpack}
\end{table}

In applying the roofline model we follow the approach suggested
in~\cite{para_pearls} and use the well-known STREAM~\cite{stream} and
LINPACK~\cite{linpack} benchmarks in order to get the upper bounds on
the memory bandwidth and FLOPS for the test hardware. Since we are
using Intel CPUs we used the Intel Math Kernel Library implementation
of LINPACK.

Interpreting the results of the STREAM benchmark in order to get
bandwidth figures for all levels of the cache hierarchy is not
straightforward. We therefore used the STREAM2 benchmark which is
intended for this purpose. In Figure~\ref{FIG_stream2} we plot the
results obtained from the DAXPY component of the STREAM2 benchmark on
the E5-2697 CPU (using v.~14 {\bf (check this)} of the Intel
compiler. The plateaux in bandwidth for L2, L3 and main memory are
clear but that for L1 is considerably noisier. The final values we
have used in the roofline plots are given in
Table~\ref{TAB_stream_linpack}.

\begin{figure}
\includegraphics[width=120mm]{stream2_archer}
\caption{Results from running the STREAM2 benchmark on the E5-2697
  CPUs. Error bars indicate the maximum and minimum values obtained
  from 200 repeats of each case.  size. }
\label{FIG_stream2}
\end{figure}

 The equivalent results for the E5-1620 CPU are shown in
 Figure~\ref{FIG_daxpy_bw}. This plot has the same shape as that for
 the E5-2697 but is shifted upwards indicating an increase in
 bandwidth accross all levels of the cache hierarchy, but particularly
 for L3 and L1. In addition, the width of the plateau corresponding to
 the L3 cache is narrower because of the smaller size of this cache on
 the E5-1620.

\begin{figure}
\includegraphics[width=120mm]{daxpy_bw_perf}
\caption{Bandwidth obtained on the E5-1620 CPU by the original and
  modified versions of the daxpy kernel from the STREAM2 benchmark.}
\label{FIG_daxpy_bw}
\end{figure}

\section{Hardware Performance Counters}

Initially, we hoped to use Hardware Performance Counters (HWPCs) in
order to get an accurate measure of the FLOPS and memory traffic
associated with specific kernels in each of the models.

The likwid tool (\url{https://github.com/RRZE-HPC/likwid}) gives
access to HWPCs on Intel and AMD CPUs.
However, there are issues with the accuracy of these counters,
especially for FLOP-related measurements on Ivy Bridge architectures. A
tool is included in the likwid distribution that tests the accuracy of
HWPC-derived data by running micro-benchmarks for which the FLOP count
and memory traffic are known:
\url{https://code.google.com/p/likwid/wiki/IvyBridgeVerification}.
The results of doing this on an Intel Xeon E5-1620 v2 CPU are shown in
Figure~\ref{FIG_flops_dp_test}. It appears that once the memory
footprint spills out of L1 cache (which all of our nemolite2d
benchmark configurations do), the performance counter results
significantly overestimate the FLOPs performed.

\begin{figure}
\centering
\includegraphics[width=120mm]{FLOPS_DP_stream_ivybridgeEP.pdf}
\includegraphics[width=120mm]{FLOPS_DP_triad_ivybridgeEP.pdf}
\caption{Results of the likwid accuracy test for the FLOPS\_DP group
  and the stream (top) and triad (bottom) benchmark on an Intel Xeon
  E5-1620 v2.}
\label{FIG_flops_dp_test}
\end{figure}

We therefore need some other way of measuring FLOPS. We could use
likwid on an older Intel CPU (to which we have root access) or try
another tool on the IvybridgeEP system. Other possible tools include
Intel's VTune or OProfile.

Despite the unreliability of the absolute values of the FLOP counts,
we hope that the trends in HWPC data as the problem size is varied
will be useful. This raw data is given in Table~\ref{TAB_hwpc_mom_u}.

\begin{table}
\begin{tabular}{l|l|l|l|l}
\hline
\hline
               & \multicolumn{4}{c}{Problem size} \\
               & 32     &    64        &    128       &   256 \\
\hline
Footprint (MB) & 0.1 & 0.5          & 2.3          & 9.2 \\
\hline
Runtime (RDTSC) [s]   & 0.144868 &  0.4485108   &   1.697142   &   6.446171  \\
Runtime unhalted [s]  & 1.199456e-01 & 4.475135e-01 & 1.750448e+00 &   6.734944e+00 \\
CPI                   & 6.431867e-01 & 6.423220e-01 & 6.503796e-01 &   6.348858e-01 \\
\hline
MFLOPS           & 3.112780e+03 & 3.877720e+03  & 4.163209e+03 & 4.304728e+03 \\
AVX MFLOPS       & 2.840625e+03 & 3.660086e+03  & 3.993397e+03 & 4.164647e+03 \\
Packed MUOPS/s   & 7.423377e+02 & 9.584935e+02  & 1.046781e+03 & 1.092280e+03 \\
Scalar MUOPS/s   & 2.077927e+02 & 1.306897e+02  & 7.294811e+01 & 3.784506e+01 \\
\hline
L1 DTLB load misses          &      2007     &      17015   &    1317912   &    6748712   \\
L1 DTLB load miss rate       &  2.949686e-06 &  6.607264e-06& 1.326630e-04 & 1.723565e-04 \\
L1 DTLB load miss durn [Cyc] &  3.970154e+01 &  3.080153e+01& 2.101039e+01 & 2.013584e+01 \\
L1 DTLB store misses         &       77      &       16     &     87257    &    401036    \\
L1 DTLB store miss rate      &  1.131668e-07 &  6.213119e-09& 8.783423e-06 & 1.024213e-05 \\
L1 DTLB store miss durn [Cyc]&  5.633766e+01 &       29     & 2.151055e+01 & 2.669793e+01 \\
\hline
L2D load bandwidth [MB/s]  &  3.811099e+03 &  5.030691e+03 & 1.080936e+04 & 1.446214e+04 \\
L2D load data volume [GB]  &   0.541220288 &   2.34655488  &  18.1215488  &   92.177536  \\
L2D evict bandwidth [MB/s] &  3.131102e+02 &  3.094459e+02 & 4.501487e+02 & 5.065637e+02 \\
L2D evict data volume [GB] &   0.04446528  &   0.144340352 &  0.75465984  &  3.22869248  \\
L2 bandwidth [MB/s]        &  4.600834e+03 &  5.493404e+03 & 1.130282e+04 & 1.498363e+04 \\
L2 data volume [GB]        &   0.653371776 &   2.562386304 & 18.948807808 & 95.501414592 \\
\hline
L3 load bandwidth [MB/s]  &  2.434255e+03 & 4.626177e+03 & 4.511686e+03 & 4.607617e+03 \\
L3 load data volume [GB]  &   0.371968512 &  2.07489024  &   7.6168768  &   29.692832  \\
L3 evict bandwidth [MB/s] &  5.652167e+02 & 4.828169e+02 & 2.771148e+02 & 2.722823e+02 \\
L3 evict data volume [GB] &   0.086368448 &  0.216548608 &  0.467840512 &  1.75466688  \\
L3 bandwidth [MB/s]       &  2.999472e+03 & 5.108994e+03 & 4.788801e+03 & 4.879899e+03 \\
L3 data volume [GB]       &   0.45833696  &  2.291438848 &  8.084717312 &  31.44749888 \\
\hline
\hline
\end{tabular}
\caption{Hardware performance counter data obtained using the likwid
  tool for the u-Momentum kernel on the Intel E5-1620 v2 processor.}
\label{TAB_hwpc_mom_u}
\end{table}

\section{Performance Analysis with the Roofline Model}

\subsection{Calculating Kernel Characteristics}

Although the Continuity kernel is very simple, the u-Momentum kernel
is not and in particular, includes a call to the {\tt sin()} function.
In order to estimate the cost of this function in FLOPs, the following
code was compiled (with the Intel compiler and -O1) and run with
likwid-perfctr:
\begin{verbatim}
  my_sum = 0.0d0
  do i = 1, 1000
    my_sum = my_sum + sin(real(i))
  end do
  write (*,*) 'My sum = ', my_sum
\end{verbatim}
This measurement revealed that the {\tt sin()} function requires
approximately 13 FLOPs.

A key component of the Roofline model is the Arithmetic Intensity of
the code being executed:
\begin{equation}
AI = \frac{\textrm{No. of floating-point operations}}{\textrm{Bytes fetched from memory}}
\end{equation}
Since HWPC data proved to be unreliable, we calculated this quantity
manually by examining the source code and counting the number of
memory references and arithmetic operations that it contained. In
doing this counting we assume that any references to adjacent array
elements are fetched in a single cache-line and thus only count once.

To illustrate this, we consider the first loop-nest in the
time-stepping section of the Shallow code. This is a doubly-nested
loop over {\tt i} and {\tt j} with {\tt i} innermost. The body of the
loop is:
\begin{verbatim}
CU(I+1,J) = .5d0*(P(I+1,J)+P(I,J))*U(I+1,J)
CV(I,J+1) = .5d0*(P(I,J+1)+P(I,J))*V(I,J+1)
Z(I+1,J+1) =(FSDX*(V(I+1,J+1)-V(I,J+1))- &
             FSDY*(U(I+1,J+1)-U(I+1,J)))/&
            (P(I,J)+P(I+1,J)+            &
             P(I+1,J+1)+P(I,J+1))
H(I,J) = P(I,J)+0.25d0*                   &
         (U(I+1,J)*U(I+1,J)+U(I,J)*U(I,J) & 
         +V(I,J+1)*V(I,J+1)+V(I,J)*V(I,J))
\end{verbatim}
This code writes to four memory locations ({\tt CU}, {\tt CV}, {\tt Z}
and {\tt H}) and reads from ten distinct locations.  However, if we
assume caching of reads from locations adjacent in memory ({\it e.g.} {\tt
  P(i,j)} and {\tt P(i+1,j)}) then we have only six distinct reads.
If we assume that each of these ten memory accesses will touch a
separate cache line then this code fragment will result in the moving
of ten cache lines, each of which is 64 bytes (eight double-precision
words). However, as noted above, this code fragment is the body of a
loop nest. Therefore, upon the next loop trip, {\tt i} will have been
incremented by one and the majority of the accessed values will still
be in cache. In fact, a new cache line will only have to be fetched
every seven trips of the (innermost) loop and thus the mean memory
traffic per loop trip is approximately $640/7$ bytes.

We can also count the number of floating-point arithmetic operations
(FLOPs) in this code fragment which gives us 12 multiplications and 12
additions. The $AI$ for this code fragment is then $24*7/640 = 0.26$
FLOPs/byte.  If we do the same analysis for the (longer) u-Momentum
kernel of nemolite2d then we find that there are 51 addition
operations, 56 multiplication/division operations and one call to {\tt
  sin()}. There are 31 distinct memory references. This gives an $AI$
of approximately $(107+13)*7/(31*64) = 0.42$ FLOPs/byte (where we
assume the call to {\tt sin()} costs 13 FLOPs). The results of this
analysis are summarised in Table~\ref{TAB_kernel_details}.

\begin{table}
\begin{tabular}{|l|c|c|c|}
\hline
                             & Shallow & \multicolumn{2}{c|}{nemolite2d} \\
WSS of time-stepping (bytes) & $104n^2$ & \multicolumn{2}{c|}{$180n^2$} \\ 
\hline
Kernel                     & 1st loop nest & Continuity & u-Momentum \\
\hline                                
FLOPs                      & 24      &    14      &   107+13   \\
Distinct memory refs       & 10      &    12      &   31       \\
$AI$ (FLOPs/byte)          & 0.26    &  0.128     &   0.42     \\
WSS (bytes)                & $56n^2$ &  $72n^2$   &   $136n^2$ \\
\hline
\end{tabular}
\caption{Details of the kernels investigated with the roofline
  model. WSS is the Working Set Size and $n$ is the problem size --
  the length of one side of the square domain.}
\label{TAB_kernel_details}
\end{table}

We measure the mean time to execute the u-Momentum kernel over the
whole domain as 0.0021 seconds (Intel v.16 on the E5-1620) for the
$256^2$ problem size. This gives us a performance of $\frac{120\times
  256^2}{0.0021} = 3.7$ GFLOPS. (In contrast, instrumenting this
section of code with the LIKWID marker API and using likwid-perfctr we
measure 4.7 GFLOPS.)  For the $128^2$ domain, the measured time is
0.556E-03 s. This gives a performance of $\frac{120\times
  128^2}{0.556\times10^{-3}} = 3.5$ GFLOPS which again is some 25\% less
than we measure directly with likwid.

Using these figures we can finally construct the Roofline model for
the two CPUs and the various kernels. In
Figure~\ref{FIG_roofline_archer} we plot the model for the E5-2697 CPU
and show the performance of the 1st loop-nest of Shallow and the
u-Momentum kernel of nemolite2d. Two things are immediately clear from
this plot: first, the performance of the first loop nest of Shallow is
clearly memory-bandwidth limited and second, the u-Momentum kernel of
NEMOLite2D is no-where near approaching this limit.

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_archer}
\caption{Comparison of the performance achieved by NEMOLite2D and
  Shallow on the E5-2697 CPU.}
\label{FIG_roofline_archer}
\end{figure}

Two performance points are plotted for the Shallow kernel
corresponding to domains of $512^2$ and $1024^2$. Using the equation
for working-set size in Table~\ref{TAB_kernel_details}, these
correspond to 14.7 and 58.7~MB, respectively. Since the E5-2697 has
an L3 cache of 30~MB, the $512^2$ case should be contained in cache
while the $1024^2$ case will spill to main memory. This is borne out
by the plot; the $1024^2$ case is achieving a performance in-line
with saturating the bandwidth to main memory. The $512^2$ case is
approaching the L3 bandwidth limit. We investigate the reasons for
the fact that it does not achieve the L3 bandwidth limit below.

In contrast to the results obtained with the Shallow kernel, the
performance of the NEMOLite2D u-Momentum kernel shows little
sensitivity to working-set size -- the $128^2$, $256^2$ and $512^2$
performance points in Figure~\ref{FIG_roofline_archer} are very close
together and well below any of the roofline limits.

In Figure~\ref{FIG_roofline_haswell} we plot the Roofline model
obtained for the E5-1620 CPU. In order to investigate the poor
performance of the u-Momentum kernel of NEMOLite2D, we also show the
performance of the (simpler) Continuity kernel from that code. As we
would expect, the performance of the u-Momentum kernel remains poor on
this CPU. For the $512^2$ domain the Continuity kernel performs well,
reaching the limit imposed by the bandwidth to main memory. However,
the working-set sizes for this kernel for the $128^2$ and $256^2$
domains are approximately 1.1~MB and 4.7~MB, respectively and
therefore both of these should fit comfortably within L3
cache. Despite this, the performance achieved with these domain sizes
falls short of the L3 bandwidth limit.  Even when we consider that the
Continuity kernel is just one of several within the time-stepping loop
of NEMOLite2D we see that the L3 bandwidth limit is still applicable;
the loop as a whole has a working-set size of $180\times n^{2}$
bytes. For the $32^2$ domain this is 180~KB ({\it i.e.}\ within L2),
while the $52^2$ (475~KB) through to the $128^2$ domain (2.9~MB) fall
well within the capacity of L3.

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_haswell}
\caption{Comparison of the performance achieved by different
  NEMOLite2D kernels on the E5-1620 CPU.}
\label{FIG_roofline_haswell}
\end{figure}

In order to investigate why the performance of the Continuity kernel
is falling well short of the maximum predicted by the roofline model
we have experimented with the level of SIMD
vectorisation. Figure~\ref{FIG_cont_vec} shows the performance of the
Continuity kernel compiled with differing levels of SIMD vectorisation
and after code modifications intended to improve vectorisation.
Compiling without auto-vectorisation (compiler flag ``-no-vec''), the
kernel achieves a performance that appears to be largely limited by
the bandwidth to main memory. Note that there is very little
opportunity for temporal cache re-use in the Continuity kernel -- only
three of the twelve distinct memory references are to previous rows of
the array. The size of this effect can be seen in the small drop in
performance (black bars) in going from the $256^2$ to the $512^2$
domain as at this point the working-set size of this kernel spills out
of L3 cache.

\begin{figure}
  \centering
  \includegraphics[width=130mm]{continuity_opt}
  \caption{Performance of the Continuity kernel with optimisations
    intended to aid SIMD vectorisation. Results are for v.16 of the
    Intel compiler on a Xeon E5-1620 v.2 CPU. Optimisations are
    cumulative for results with a given SIMD instruction set.}
  \label{FIG_cont_vec}
\end{figure}

The E5-1620 CPU supports two different SIMD instruction sets;
Streaming SIMD Extensions (SSE; flag ``-axSSE4.2'') and Advanced
Vector Extensions (AVX; flag ``-xHost''). SSE supports vector lengths
of 128 bits (two double-precision words) and AVX 256 bits (four
double-precision words). For the problem sizes that largely fit within
L3 or L2 cache, switching on the use of SSE instructions provides a
speed-up of between 1.5 and 1.8 - the factor increasing with problem
size until the WSS of the time-stepping loop begins to fall out of L3
cache. It is therefore likely that the reason this factor is below the
ideal value of 2.0 is the overhead associated with constructing the
loop and handling any ``peel'' or ``remainder'' loops. As the problem
size increases, this overhead (as a fraction of the vectorisable
workload) decreases.

Examining the optimisation report produced by the Intel compiler
revealed that the majority of array accesses within the loop were not
aligned:
\begin{verbatim}
Example of unaligned access report here
\end{verbatim}
and that this inhibited efficient vectorisation. Tackling this issue
requires two steps; first, ensure that the arrays are allocated to
align with 64-byte boundaries and second, ensure that the compiler is
aware of this where those arrays are used. The first requirement is
easily met by passing the ``-align array64byte'' flag to
the compiler when building the code. This ensures that any array that
is constructed using a call to ALLOCATE is aligned on 64-byte
boundaries in memory.

The second requirement is complicated by the fact that the NEMOLite2D
code makes extensive use of Fortran derived types --- the data arrays
used for computation are members of derived types. The information
about the alignment of these arrays cannot therefore be propagated
through the code by module {\tt use} statements. In fact, even the
Intel compiler directive which can be used to specify the alignment
of a particular array at its point of use will not accept references
to members of a derived type. It was therefore necessary to wrap the
kernel call (and associated loops) within a routine and pass in
references to the arrays. Within the wrapper routine the corresponding
dummy arguments are declared as simple arrays and their alignment can
then be specified using the {\tt ASSUME\_ALIGNED} directive:
\begin{verbatim}
subroutine invoke_continuity_arrays(nx, ny, ssha, ...)
  integer, intent(in)   :: nx, ny
  real(wp), intent(out) :: ssha(nx,ny)
!DIR$ ASSUME_ALIGNED ssha:64
  ...
\end{verbatim}
This approach also enables the extent of the arrays ({\tt nx} and {\tt
  ny} in this example) to be specified explicitly.

The continuity kernel includes array accesses of the form {\tt
  a(i,j-1)}. If array {\tt a} is allocated with extent {\tt (Nx, Ny)}
then this specifies a location in memory {\tt Nx} words before that
occupied by {\tt a(i,j)}. In order for such an access to be aligned,
{\tt Nx} must be a multiple of the SIMD vector width (2 for SSE, 4 for
AVX) and this must be specified to the compiler. We therefore padded Nx
as required and added a directive within the wrapper routine:
\begin{verbatim}
!DIR$ ASSUME (MOD(Nx,ALIGNMENT) .EQ. 0)
\end{verbatim}
where {\tt ALIGNMENT} is an integer parameter declared to have the value two
when compiling with SSE and four with AVX.

We can see in Figure~\ref{FIG_cont_vec} that these changes (``SSE
align'') significantly improve the performance of the smaller
domains. Again, as the domain size increases this improvement reduces
until, for $512^2$ and $1024^2$ (where we are limited by the bandwidth
to main memory), this optimisation has no effect.

Since the Continuity kernel is one of several in the time-stepping
loop, we experimented with pre-loading the fields into cache simply by
executing the Continuity kernel immediately prior to executing it for
a second time with timing. Figure~\ref{FIG_cont_vec} shows that this
``pre-loading'' slightly improved the kernel performance for those
problem sizes that fit within cache. {\bf Discuss this}

Once all of the above steps were complete, the only un-aligned array
accesses reported by the compiler were those of the form {\tt
  a(i-1,j)}. Without taking a physical copy of the array {\tt a}, such
unaligned accesses are unavoidable in a stencil-based
finite-difference code. However, in order to quantify how much these
accesses limit the performance of the kernel we experimented with
replacing e.g. {\tt a(i-1,j)} with {\tt a(i-ALIGNMENT,j)}. The
resulting kernel had no un-aligned accesses and performed
significantly better; actually reaching or exceeding the L3-bandwidth
roofline for domains $64^2$ -- $256^2$.

The final SIMD-related optimisations are concerned with what happens
at the beginning and end of a SIMD-vectorisable loop. If the loop
updates e.g. {\tt a(i,j)} and begins on a value of {\tt i} that
  results in unaligned memory accesses (e.g. {\tt i = 2}) then the
  compiler must insert a ``peel'' loop to handle those iterations
  required to reach a point where memory is accessed on a 64-byte
  boundary.  Similarly, if the trip-count of the resulting
  vectorisable loop is not an integer multiple of the SIMD vector
  width then the compiler must insert a ``remainder'' loop to handle
  those iterations left over.  We removed the remainder loop by
  ensuring that the length of the inner loop was a multiple of the
  vector width and adding an {\tt ASSUME} compiler directive
  (c.f. that for {\tt Nx} above). This was straightforward since we
  had already padded the array extents themselves and NEMOLite2D uses
  a mask to define the active computational domain - the additional
  grid points resulting from the padding were simply marked as
  external to the domain.  Despite doing this we found that the
  remainder loop was not removed until we manually introduced the peel
  loop (rather than relying on the compiler).

The inner loop of the kernel where all array accesses are aligned and
the number of distinct streams has been reduced to eight (with a
concommitant reduction in the number of FLOPs):
\begin{verbatim}
do ji = ALIGNMENT+1, M, 1
   rtmp1 = (sshn_u(ji  ,jj ) + hu(ji  ,jj  ))*un(ji  ,jj)
   rtmp2 = (sshn_u(ji-ALIGNMENT,jj ) + hu(ji-ALIGNMENT,jj))* &
           un(ji-ALIGNMENT,jj)
   rtmp3 = (sshn_v(ji ,jj )  + hv(ji  ,jj  ))*vn(ji ,jj)
   ssha(ji,jj) =  (rtmp2 - rtmp1 - rtmp3) / sshn_t(ji,jj)
end do
\end{verbatim}

Figure~\ref{FIG_kernel_perf} shows the performance of the Continuity
and u-Momentum kernels when compiled without vectorisation
(``-no-vec''), with SSE vectorisation (``-axSSE4.2'') and with AVX
(``-xHost''). SSE supports vector lengths of 128 bits (two
double-precision words) and AVX 512 bits (four double-precision
words). For the 64, 128 and 256 problem sizes, changing from no
vectorisation to SSE increases the performance of the Continuity
kernel by up to roughly 60\%; some way short of the ideal speed-up of
100\% that would be obtained with perfect SSE vectorisation of a
compute-bound kernel. Unsurprisingly then, moving from SSE to AVX
vectors does not give any performance improvement and in fact, is
slightly detrimental for most of the problem sizes.

\begin{figure}
  \centering
  \includegraphics[width=120mm]{nemolite2d_kernel_perf}
  \caption{Performance of the Continuity and u-Momentum kernels as a
    function of SIMD vectorisation and problem size. Results are for
    v.16 of the Intel compiler on a Xeon E5-1620 v.2 CPU.{\bf This plot
requires updating or dropping in favour of the clustered-bar form.}}
  \label{FIG_kernel_perf}
\end{figure}

Peak absolute performance is obtained for the $256^2$ problem size
with SSE instructions. The 512 problem size spills from L3 cache and
becomes bound by the bandwidth to main memory: SIMD vectorisation
gives no performance benefit. The fact that best performance is
obtained when the problem size fits within L3 indicates that there are
other overheads that dominate when the problem size is small enough to
fit within L2. We checked the efficiency of the SIMD vectorised code
by looking at the compiler diagnostic output. On the basis of this we
added Intel-specific directives to the code in order to specify that
the arrays were aligned on 64-byte boundaries. We also manually coded
the 'peel loop' (iterations required to reach a 64-byte
boundary). Having done this the compiler reported that no peel or
remainder loops were generated and the resulting (SSE) vectorised loop
had an estimated speed-up of 1.9. (This estimated speed-up was
previously 1.6; in agreement with the speed-up we measured.) After
making these changes the performance of the Continuity kernel with the
$32^2$ domain increased from 4.22 to 4.57 GFLOPs.

Turning now to the u-Momentum kernel (lower plot in
Figure~\ref{FIG_kernel_perf}), it is striking that the kernel
performance is almost independent of problem size. The reason for this
becomes clear when we edit the kernel to remove the many IF statements
(present because of the application of boundary conditions). The blue
symbols mark the resulting performance which is significantly
increased and again reaches a maximum at the 256 problem size.  It
should be noted that the compiler reported that this loop had been
vectorised, even before the IF statements were removed. Clearly the
vectorisation is more efficient once the code is simplified. However,
even with the up-lift in performance given by removing all the IF
statements, the performance of this kernel is still far short of even
the main-memory bandwidth limit, despite the fact that the $256^2$
problem fits within L3 cache.

In order to help further understand the performance of the Continuity
kernel in Figure~\ref{FIG_cont_perf} we plot the performance of the
Continuity kernel alongside HWPC-derived metrics (obtained using the
likwid tool). We use the version compiled with SSE instructions since
that gave the best performance. This plot confirms that the $32^2$
problem size is dominated by bandwidth to L2. Sizes $64^2$, $128^2$
and $256^2$ are dominated by bandwidth to L2 and L3 and $512^2$ makes
heavy use of main memory.

\begin{figure}
  \centering
  \includegraphics[width=120mm]{nemolite2d_cont_perf}
  \caption{Performance of the Continuity kernel (open circles -- right y-axis)
    and associated HWPC data (bars -- left y-axis) as a function of
    problem size. Results are for v.16 of the Intel compiler on a Xeon
    E5-1620 v.2 CPU.}
  \label{FIG_cont_perf}
\end{figure}

As well as examining the performance of the Continuity kernel, we have
also experimented with altering the DAXPY kernel of the STREAM2
benchmark so as to vary its $AI$. We did this by adding further
arithmetic operations but making use of the same array
elements. We used five different kernels:
\begin{enumerate}
\item {\tt y(i) = y(i) + a*x(i)}, AI = 0.125
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)}, AI = 0.167
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)}, AI = 0.208
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)*y(i)}, AI = 0.25
\item {\tt y(i) = y(i) + a*x(i)+ x(i-1)*y(i)*y(i)*y(i)}, AI = 0.25
\end{enumerate}
The fifth of these introduces an un-aligned array access ({\tt
  x(i-1)}) so as to better represent real-life finite difference
kernels.  Figure~\ref{FIG_daxpy_roofline} shows the performance of
these kernels on a roofline plot for the E5-1620 CPU. The performance
of each kernel is plotted for four problem sizes, chosen so as to fit
within L1, L2, L3 cache and main memory. The performance of kernels 2,
3 and 4 is entirely consistent with the roofline model for problem
sizes that fit within the L3 and L3 caches (red and black bars). Their
performance for the smallest problem size doesn't quite reach the L1
bandwidth limit but does appear to track the roofline.  Kernel 1 has
the lowest AI of the five kernels and this appears to limit its
performance; at no point does the performance reach the limit imposed
by the bandwidth to the relevant level of the memory hierarchy. {\bf need to expore the reason for this}

The performance of kernel 5 is plotted using patterned bars in
Figure~\ref{FIG_daxpy_roofline}. This kernel has the same AI as kernel
4 but has an un-aligned array access. Comparing the performance of
these two kernels we see that the resulting reduction in the
efficiency with which the kernel can be SIMD vectorised (because of
the need for gather/scatter operations) harms performance,
particularly when the cost of accessing memory is small (i.e. when the
problem is contained in L1 or L2). Since all of the NEMOLite2D kernels
contain un-aligned array accesses, this goes some way to accounting
for the fact that they do not reach the theoretical limits of the
roofline plot in Figure~\ref{FIG_roofline_haswell}.

\begin{figure}
\includegraphics[width=120mm]{roofline_daxpy}
\caption{Roofline plot of daxpy-based kernels on the E5-1620 CPU. Each
  cluster of bars represents a different kernel. The green, black, red and
  blue bars are for problem sizes chosen to fit within main memory, L3, L2 and
  L1 caches (arrays of 3.3M, 94K, 8.3K and 1.5K elements),
  respectively. Bars with patterned fill are for kernel 5 (unaligned
  access).}
\label{FIG_daxpy_roofline}
\end{figure}

\section{Tools}

This work has deliberately focused on the application of a relatively
simple performance model to two very simple, stencil based simulation
codes.  Despite this, getting the data necessary for the construction
of the roofline model has not been straightforward. Once the roofline
model had been constructed, it was also difficult to establish why the
performance of a section of code was apparently poor.

The reasons for this include:
\begin{itemize}
\item HWPC data on its own doesn't give much insight into performance issues;
\item FLOPS data produced by likwid is unreliable on Ivy Bridge processors;
\item Persuading any tool to give HWPC data can be difficult. likwid
  is relatively straightforward but root access was required on a
  Redhat-based linux distribution. Thus far we haven't succeeded in
  getting this data from Intel's VTune.
\item The Intel compiler tends to report ``Loop was vectorised'' but
  it is hard to judge how efficient this vectorisation is. For
  instance we have found occasions where backing-off from AVX to SSE
  instructions actually gave better performing code. Ways of
  establishing the quality of the code produced by the compiler are
  missing;
\end{itemize}

It feels as though there is a need for a tool that addresses some of
these issues. It could be that Intel's Vectorization Advisor will be
useful and this requires investigation.

\section{Conclusions}

\bibliographystyle{unsrt}
\bibliography{gocean_roofline}

\end{document}