### 2016-11-15 docs/papers/roofline_habakkuk/habakkuk.tex
Beginnings of skeleton of Habakkuk paper

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage{url}

\newcommand{\psykal}{{PS}y{KA}l}

\begin{document}

\title{Analysing the Performance of Shallow-Water Models with the
 Roofline Model}

\author{A.~.R.~Porter and R.~W.~Ford}

\maketitle

\section{Introduction}

Shallow implementation of Sadourny~\cite{sadourny75}.

\section{The Roofline Model and its Limitations}

\subsection{What the Roofline Model (RFM) is}

Arithmetic intensity and MAX(memory bandwidth, FLOPS)

\subsection{Approaches to constructing a RFM}

Theoretical limits versus experimentally-measured values (STREAM and
LINPACK).

Measuring Arithmetic Intensity. Hardware Performance Counters (access,
reliability, difficulty of interpretation) or eye-balling the code
(error prone and time consuming).

\subsection{Basic Assumptions}
\subsection{When the assumptions break down}

All FLOPs are not equal. For instance, on Ivy Bridge a division costs
approximately eight times as much as a multiplication.
Instruction-level parallelism - operations despatched to different
execution ports can be executed in parallel.

\section{How to remove some of the assumptions}

Parse the (Fortran) source code.

\section{Conclusions}

\bibliographystyle{unsrt}
\bibliography{habakkuk}

\end{document}