### 2017-06-23 docs/papers/nemolite2d_performance/nemolite2d_perf_abstract.tex
nemolite-paper: add separate abstract tex file

%% 2-column papers and discussion papers
\documentclass[gmd, manuscript]{copernicus}

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\newlength{\picwidth}
\setlength{\picwidth}{83mm}

\begin{document}

\title{Portable Multi- and Many-Core Performance for Finite Difference Codes;
  Application to the Free-Surface Component of NEMO.}

\Author[1]{Andrew}{Porter}
\Author[2]{Jeremy}{Appleyard}
\Author[1]{Mike}{Ashworth}
\Author[1]{Rupert}{Ford}
\Author[3]{Jason}{Holt}
\Author[3]{Hedong}{Liu}
\Author[4]{Graham}{Riley}

\affil[1]{Science and Technology Facilities Council, Daresbury Laboratory, UK}
\affil[2]{NVIDIA Corporation}
\affil[3]{National Oceanography Centre, Liverpool, UK}
\affil[4]{University of Manchester, Manchester, UK}

\runningtitle{Portable Multi- and Many-Core Performance for Finite Difference Codes}

\runningauthor{A.~R.~Porter et al}

\correspondence{Andrew Porter (andrew.porter@stfc.ac.uk)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

\firstpage{1}
\maketitle

\begin{abstract}

We present an approach which we call {PS}y{KA}l that is designed to achieve
portable performance for parallel, finite-difference Ocean models.  In
{PS}y{KA}l the code related to the underlying science is formally
separated from code related to parallelisation and single-core
optimisations. This separation of concerns allows scientists to code
their science independently of the underlying hardware architecture
and for optimisation specialists to be able to tailor the code for a
particular machine independently of the science code. We have taken
the free-surface part of the NEMO ocean model and created a new,
shallow-water model named NEMOLite2D. In doing this we have a code
which is of a manageable size and yet which incorporates elements of
full ocean models (input/output, boundary conditions, \textit{etc.}).
We have then manually constructed a {PS}y{KA}l version of this code and
investigated the transformations that must be applied to the
middle/PSy layer in order to achieve good performance, both serial and
parallel. We have produced versions of the PSy layer parallelised with
both OpenMP and OpenACC; in both cases we were able to leave the
natural-science parts of the code unchanged while achieving good
performance on both multi-core CPUs and GPUs. In quantifying whether
or not the obtained performance is `good' we also consider the
limitations of the basic roofline model and improve on it
by generating kernel-specific CPU ceilings.

\end{abstract}

\end{document}