### 2017-09-14 docs/papers/nemolite2d_performance/referee_response.txt
nemolite-paper: add response to first referee

The authors thank the reviewers for their detailed feedback on our paper.
We address the points raised below, beginning with the first (anonymous)
referee and then moving on to those from Carlos Osuna.

Anonymous referee:

1/ The reviewer is right to be concerned with seeing how much
   performance may be lost by the re-structuring required for
   PSyKAl. However, the original form of the NEMOLite2D code that we
   started with for this work was unoptimised. This is emphasised by
   the results in Figure 4 which show that, for the majority of cases,
   the final PSyKAl version was significantly faster. In contrast, our
   previous paper (Porter et al, 2016) tackled the "Shallow" code
   which has had optimisations applied to it over a period of some
   twenty years and was therefore a much more demanding test. There,
   the initial PSyKAl re-structuring did significantly harm
   performance. However, we showed that it was possible to recover the
   performance of the optimised original by applying code
   transformations that were PSyKAl-compliant. We therefore do not
   think that showing results for the optimisations
   applied to the original form of NEMOLite2D is necessary.

2/ The PSyKAl approach is in fact applicable to both Finite Element and
   Finite Difference codes (and in fact, LFRic, the largest project making use
   of it is using Finite Elements). We have updated the title to make
   this clear.

3/ A new section, "Related Approaches", has been added to the paper. In this
   we bring together a comparison of various other approaches including OCCA.
   This makes explicit the differences and makes clearer the advantages of
   PSyKAl.

4/ We also use the new "Related Approaches" section to emphasise that
    although the tool to generate the PSy layer exists ("PSyclone"),
    the current paper is about code structuring.