### 2017-10-11 docs/papers/nemolite2d_performance/referee_response.txt
nemolite2d-paper: update referee response to match copernicus requirements

A. Kerkweg:
==========

We have added the name of the model used/developed in this work to the title.

This model has been published with a DOI and a reference to this has been
added to the Code Availability section of the paper.

Anonymous referee:
=================

"1. The optimization and performance sections are very well done,
however, I would like to see one additional data point in the parallel
performance sec- tion: all of the optimizations and directives applied
to the original NEMOLite2D code.  This will allow the reader to see if
much or any performance is being left on the table by using the PSyKAl
approach"

Our response:

1/ The reviewer is right to be concerned with seeing how much
   performance may be lost by the re-structuring required for
   PSyKAl. However, the original form of the NEMOLite2D code that we
   started with for this work was unoptimised. This is emphasised by
   the results in Figure 4 which show that, for the majority of cases,
   the final PSyKAl version was significantly faster. In contrast, our
   previous paper (Porter et al, 2016) tackled the "Shallow" code
   which has had optimisations applied to it over a period of some
   twenty years and was therefore a much more demanding test. There,
   the initial PSyKAl re-structuring did significantly harm
   performance. However, we showed that it was possible to recover the
   performance of the optimised original by applying code
   transformations that were PSyKAl-compliant. We therefore do not
   think that showing results for the optimisations
   applied to the original form of NEMOLite2D is necessary.

--

"2.  The title of the paper suggests the PSyKAl approach is limited to
finite-difference models, yet there is barely any mention of this
limitation in the paper.  Can you describe what causes this limitation
or if there will be attempts to extend the idea to more complex
finite-element or finite-volume models?"

Our response:

2/ The PSyKAl approach is in fact applicable to both Finite Element and
   Finite Difference codes (and in fact, LFRic, the largest project making use
   of it is using Finite Elements). We have updated the title to make
   this clear.

--

"3.  I fully agree that a separation of concerns is necessary to
achieve portable performance, however it is not clear to me how
significantly different PSyKAl is from similar ideas, such as OCCA
(http://libocca.org) where kernels are written in a simple C or
Fortran based kernel language and translated into OpenMP, CUDA or
OpenCL and invoked similarly to a CUDA kernel.  Also, time-stepping
research has long assumed that the spatial domain is computed using a
single function call which iterates over all of points and executes
the appropriate kernels.  Could the authors be more specific about the
contributions which the PSyKAl approach makes, citing specific
differences from ex- isting programming models, like CUDA and OpenMP?"

Our response:

3/ A new section, "Related Approaches", has been added to the paper. In this
   we bring together a comparison of various other approaches including OCCA.
   This makes explicit the differences and makes clearer the advantages of
   PSyKAl.

--

"4.  The paper briefly mentions that generation of the PSy layer will
be automated in the future.  Perhaps the authors could emphasize this
more (as part of any changes made #3), as this paper describes a
number of lessons learned which will be applied to work on automation."

Our response:

4/ We also use the new "Related Approaches" section to emphasise that
    although the tool to generate the PSy layer exists ("PSyclone"),
    the current paper is about code structuring.

Carlos Osuna's general review:
=============================

"While the mathematics and equations are well described, the reviewer
is missing some snippet of code that is representative of the target
model.  Showing some simplified codes, like the dominant momentum
section, could help the reader to understand the computational
patterns (see some floating point operations performed, extensively
dis- cussed in section 3.3), and, in general, better reason about
sections that discuss the relevant characteristics of the code, like
AI, or ILP."

Our response:

We have added a few key lines of source code from the Momentum kernel
in section 3.3. We have also added text that explains that the
apparently poor performance of this kernel is down to the number of
division operations.  On the Ivy Bridge, division operations require
at least eight times as many cycles as a multiplication which can be
very significant.

--

"A small "Related Work" section where other ap- proaches are explained
and compared would help and can be used to highlight the novelties of
this work. Similarly a section or paragraph describing the limitations
would be valuable. Particularly a discussion on the design choices of
PSyKAI in three levels, where all the computing architecture dependent
optimizations happen in the PSy layer.  Is it the right choice, could
the author imagine other optimizations that would need to be applied
to the Kernel layer?"

Our response:

The suggestion of a "Related Work" section is a good one and we have added
this just after the Introduction.

--

"What is there in the implementation, in terms of floating point
operations that makes this code perform badly?  Also in the comparison
of SSE and AVX, it is said that SSE and AVX versions of the division
vector instruction does not provide any benefit. That would explain
why they do not perform well, but not why SSE performs better than
AVX."

Our response:

We have added text in section 3.3. to explain that the poor performance
is due to liberal use of the division operation.

We attribute the better performance of the SSE version of the code to the
fact that the poor SIMD efficiency of the code is failing to amortize the
larger cost of the peel and remainder loops associated with the greater
vector width of AVX. We have expanded upon this in the text.

Carlos Osuna's specific comments:
================================

"pag.  7, line 10: "In order that we could safely add such directives
we altered the GOcean infrastructure to allocate all field-data arrays
with extents greater than strictly required": It is not clear to the
reader why these larger extents are required in order to add this
directive."

Our response:

(Top of page 8.) We have clarified why allocating larger arrays was
necessary in order to use the safe_address directive.

--

"pag. 7, line 22: "Although the Intel compiler does do" I would
replace by "Although the Intel compiler can do" since it is not
guarantee when the compiler will inline."

Our response:

Changed "Intel does..." to "Intel can..." as requested.

--

"section 2.1.5:  the discussion makes the reader wonder whether attaching the
grid to the field, which can be read only or not is the right choice."

Our response:

Sec 2.1.5 - the grid is not attached to the field, rather each field
keeps a pointer back to the grid upon which it is defined. This approach
was chosen in order to support applications which use more than one grid.

--

"section 2.1.8: it is not clear why inlining these copies gives better
performance. Is it due to data reuse of the same fields that might be
in cache? If so, that will depend on the domain size? Maybe a sentence
clarifying that would help."

Our response:

Sec 2.1.8 - it is not clear why in-lining manually helps (especially since
one would expect the Intel compiler in particular to do this). We suspect
that the change in the PSy-layer code causes the compiler to make some
different optimisation decisions and these happen to be beneficial. A
proper investigation of this is outside the scope of this paper. We have
added a sentence about this.

--

"section 2.2.1: Aside note: It looks like other omp strategies like
blocking would help increasing the data locality of computations among
cores computing different blocks, or parallel do simd that would allow
collapsing the loops and increase parallelism?"

Our response:

Sec 2.2.1 - yes, other strategies might be beneficial, especially blocking.
The key point is that adopting the PSyKAl structure does not limit these
possibilities.

--

"Figure 3: why is gnu still much slower?  We would expect that PSyKAI
would help the compiler adding the corresponding directives so that
all compilers could deliver similar performance. Still is a factor 2x
away from others. Can the reason for this be clarified?"

Our response:

Figure 3 - The Gnu results are much slower because gfortran is only SIMD
vectorising the loop containing the Continuity kernel. For some reason
it does not vectorise any others. Unlike Cray, there is no directive
that can be inserted to mandate that a loop be SIMD vectorised. We have
added text to this effect.

--

"page 15, line 5: Text says that 256ˆ2 fits in cache.  Which cache are
the authors referring to? A single field with that domain size would
take 512 KB (in double precision) which will be out of L1"

Our response:

Page 15, line 5 - we were referring to the L3 cache. This has been clarified.

--

"Figure 5: same comment I made before.  It seems that we can not
attain same perf.  with all compilers. Would be interesting clarify
what is the intrinsic reason for this, if this is a real limitation of
the compiler, or future work can improve the gnu performance."

Our response:

Figure 5 - it is a reflection of the fact that Gnu is free while the Intel and
Cray compilers have performance of the generated code as a key part of their
respective business models. We have added text about the lack of a crucial
directive to force the Gnu compiler to SIMD vectorise a given loop.

--

"page 18, line 9: text says that 32x32, due to limited amount of
parallelism inhibits scaling, therefore focus is on 256x256 domain
sizes.  Is it clear that is parallelism what limits scalability? The
curve shown by 32x32 might have a reasonable shape for memory bound
codes, where with few cores one can saturate the memory bandwidth of
the Ivy Bridge, and adding more cores might not provide better
performance. On the contrary the large gap between 32x32 and the other
domains suggests that the code is not cache oblivious, where the
updated points/s would not (so strongly) depend on the domain size?
See for example: https://arxiv.org/pdf/1410.5010.pdf"

Our response:

Page 18, line 9 - the NEMOLite2D performance for the 32x32 domain is
significantly greater than for the other domains because it largely
fits within L2 cache. We have calculated that the working-set size for
the momentum kernel is 136*n^2. For 32x32 this gives ~136KB while for
64x64 this gives ~544KB. The former will fit within L2 cache while the
latter will spill to L3 cache. In addition, recall that we only parallelise
the outermost loop (Sec. 2.2.1) so, for the 32x32 domain, there are only
32 items to be shared amongst the threads. We have added a note to the
text regarding this.

--

"Figure 11: For domain size of 64, parallelism will be very small on
the GPU, if col- lapse(2) is not used.  If used, 4096 GPU threads
could be active, still not 100% occu- pancy but much better. So it is
not clear why the collapse(2) didnt help in performance?"

Our response:

Figure 11 - the size of the Momentum kernels meant that
their resource use limited the occupancy that could be achieved.
Therefore COLLAPSE might only help the performance of the other kernels and
they only account for some 30% of the run-time. In addition, the kernels
had been parallelised by enclosing them within a !$acc kernels region
which leaves the details to the compiler. It would seem that in this
case the decisions made by the compiler were good and COLLAPSE provided
little or no benefit. The text has been altered to cover both of these
points.

--

"Page 22, line 11: The statement says that the comparison between
PSyKAI approach and the CUDA fortran implementation proves little
overhead introduced by PSyKAI.  While this is true, it is no proven
that the implementation is efficient, since also a CUDA implementation
could be inefficient.  Figure 12 suggests that actually the GPU imple-
mentation provides an expected performance in comparison with memory
bandwidths of the IB and K40. Adding the GPU numbers to the Roofline
model of Figure 13 would help in proving efficency of the GPU code and
make a more solid statements about the GPU performance."

Our response:

Page 22, line 11 - we agree with the reviewer's point that there was
no proof that the CUDA version was performing well. However, it is
also not a matter of comparing memory bandwidths on the CPU and GPU -
our work on the roofline model has demonstrated that the Momentum
kernel is not memory-bandwidth bound. Instead, we have added a
paragraph that demonstrates we have used profiling to investigate the
performance limitations on the GPU and are confident that we can do no
better without changing kernel code.

--

"Figure 13: font size and lines are small and hard to read. Increasing
the font size and probably using markers would help"

Our response:

Figure 13 - font size and line widths have been increased. Labels have been
re-arranged to improve clarity.